---
title: "MITx analytics edge 15.071x - UNIT 5 "
output: html_notebook
---
#Detecting Vandalism on Wikipedia

Wikipedia is a free online encyclopedia that anyone can edit and contribute to. It is available in many languages and is growing all the time. On the English language version of Wikipedia:

There are currently 4.7 million pages.
There have been a total over 760 million edits (also called revisions) over its lifetime.
There are approximately 130,000 edits per day.
One of the consequences of being editable by anyone is that some people vandalize pages. This can take the form of removing content, adding promotional or inappropriate content, or more subtle shifts that change the meaning of the article. With this many articles and edits per day it is difficult for humans to detect all instances of vandalism and revert (undo) them. As a result, Wikipedia uses bots - computer programs that automatically revert edits that look like vandalism. In this assignment we will attempt to develop a vandalism detector that uses machine learning to distinguish between a valid edit and vandalism.

The data for this problem is based on the revision history of the page Language. Wikipedia provides a history for each page that consists of the state of the page at each revision. Rather than manually considering each revision, a script was run that checked whether edits stayed or were reverted. If a change was eventually reverted then that revision is marked as vandalism. This may result in some misclassifications, but the script performs well enough for our needs.

As a result of this preprocessing, some common processing tasks have already been done, including lower-casing and punctuation removal. The columns in the dataset are:

* Vandal = 1 if this edit was vandalism, 0 if not.
* Minor = 1 if the user marked this edit as a "minor edit", 0 if not.
* Loggedin = 1 if the user made this edit while using a Wikipedia account, 0 if they did not.
* Added = The unique words added.
* Removed = The unique words removed.
Notice the repeated use of unique. The data we have available is not the traditional bag of words - rather it is the set of words that were removed or added. For example, if a word was removed multiple times in a revision it will only appear one time in the "Removed" column.

_______________________________________________________________________________________________________________

######Problem 1.1 - Bags of Words

Load the data wiki.csv with the option stringsAsFactors=FALSE, calling the data frame "wiki". Convert the "Vandal" column to a factor using the command wiki$Vandal = as.factor(wiki$Vandal).

How many cases of vandalism were detected in the history of this page?
```{r}
wiki = read.csv("Downloads/wiki.csv", stringsAsFactors = FALSE)
str(wiki)
wiki$Vandal = as.factor(wiki$Vandal)
str(wiki)
```
>1815

######Problem 1.2 - Bag of Words

We will now use the bag of words approach to build a model. We have two columns of textual data, with different meanings. For example, adding rude words has a different meaning to removing rude words. We'll start like we did in class by building a document term matrix from the Added column. The text already is lowercase and stripped of punctuation. So to pre-process the data, just complete the following four steps:

1) Create the corpus for the Added column, and call it "corpusAdded".
2) Remove the English-language stopwords.
3) Stem the words.
4) Build the DocumentTermMatrix, and call it dtmAdded.

If the code length(stopwords("english")) does not return 174 for you, then please run the line of code in this file, which will store the standard stop words in a variable called sw. When removing stop words, use tm_map(corpusAdded, removeWords, sw) instead of tm_map(corpusAdded, removeWords, stopwords("english")).

How many terms appear in dtmAdded?
```{r}
corpusAdded= Corpus(VectorSource(wiki$Added))
corpusAdded[[1]]$content
corpusAdded = tm_map(corpusAdded, removeWords, stopwords("english"))
corpusAdded = tm_map(corpusAdded,stemDocument)
```
> terms : 6675

######Problem 1.3 - Bag of Words 

Filter out sparse terms by keeping only terms that appear in 0.3% or more of the revisions, and call the new matrix sparseAdded. How many terms appear in sparseAdded?

```{r}
dtmAdded = DocumentTermMatrix(corpusAdded)
sparseAdded = removeSparseTerms(dtmAdded, 0.997)
sparseAdded
```
> terms : 166

######Problem 1.4 Bag of Words

Convert sparseAdded to a data frame called wordsAdded, and then prepend all the words with the letter A, by using the command:
**bold**colnames(wordsAdded) = paste("A", colnames(wordsAdded))**bold**

Now repeat all of the steps we've done so far (create a corpus, remove stop words, stem the document, create a sparse document term matrix, and convert it to a data frame) to create a Removed bag-of-words dataframe, called wordsRemoved, except this time, prepend all of the words with the letter R:
**bold**colnames(wordsRemoved) = paste("R", colnames(wordsRemoved))**bold**
How many words are in the wordsRemoved data frame?

```{r}
wordsAdded= as.data.frame(as.matrix(sparseAdded))
colnames(wordsAdded) = paste("A", colnames(wordsAdded)) # prepend all the wordswith the letter "A"


corpusRemoved= Corpus(VectorSource(wiki$Removed))

corpusRemoved=tm_map(corpusRemoved, removeWords, stopwords("english"))
corpusRemoved=tm_map(corpusRemoved, stemDocument)

dtmRemoved = DocumentTermMatrix(corpusRemoved)
sparseRemoved = removeSparseTerms(dtmRemoved, 0.997)
sparseRemoved

wordsRemoved= as.data.frame(as.matrix(sparseRemoved))
```
######Problem 1.5 Bag of Words
Combine the two data frames into a data frame called wikiWords with the following line of code:

**bold**wikiWords = cbind(wordsAdded, wordsRemoved)**bold**

The cbind function combines two sets of variables for the same observations into one data frame. Then add the Vandal column (HINT: remember how we added the dependent variable back into our data frame in the Twitter lecture). Set the random seed to 123 and then split the data set using sample.split from the "caTools" package to put 70% in the training set.

What is the accuracy on the test set of a baseline method that always predicts "not vandalism" (the most frequent outcome)?

```{r}

wikiWords = cbind(wordsAdded, wordsRemoved)
wikiWords$Vandal = wiki$Vandal

set.seed(123)
spl = sample.split(wikiWords$Vandal, 0.7)

train=subset(wikiWords, spl == TRUE)
test=subset(wikiWords, spl == FALSE)


table(test$Vandal) #baseline accuracy for "not vandalism"
```
>618/(618+545) = 0.5313844

######Problem 1.5 Bag of Words
Build a CART model to predict Vandal, using all of the other variables as independent variables. Use the training set to build the model and the default parameters (don't set values for minbucket or cp).

What is the accuracy of the model on the test set, using a threshold of 0.5? (Remember that if you add the argument type="class" when making predictions, the output of predict will automatically use a threshold of 0.5.)
```{r}
#cart model
wikiCart = rpart(Vandal~., data = train, method = "class")
prp(wikiCart)

#accuracy of our model on the test set, i.e. out-of-sample. 
pred = predict(wikiCart, newdata=test, type = "class")
prediction = table(test$Vandal, pred)
```
> Accuracy is (614+19)/(4+614+526+19) = 0.544282

_______________________________________________________________________________________________________________
######Problem 2.1 - Problem-specific Knowledge

```{r}
grepl("cat","dogs and cats",fixed=TRUE)
grepl("cat","dogs and rats",fixed=TRUE)

wikiWords2 = wikiWords

wikiWords2$HTTP = ifelse(grepl("http",wiki$Added,fixed=TRUE), 1, 0)

table(wikiWords2$HTTP)
```
######Problem 2.2 - Problem-specific Knowledge
```{r}
wikiCART2 = rpart(Vandal~., data=wikiTrain2, method="class")
testPredictCART2= predict(wikiCART2, newdata=wikiTest2,type="class"
table(wikiTest2$Vandal, testPredictCART2)
```
> Accuracy is (605+64)/(605+13+481+64) = 0.5752365

#######Problem 2.3 - Problem-specific Knowledge
Another possibility is that the number of words added and removed is predictive, perhaps more so than the actual words themselves. We already have a word count available in the form of the document-term matrices (DTMs).
Sum the rows of dtmAdded and dtmRemoved and add them as new variables in your data frame wikiWords2 (called NumWordsAdded and NumWordsRemoved) by using the following commands:
**bold**wikiWords2$NumWordsAdded = rowSums(as.matrix(dtmAdded))**bold**
**bold**wikiWords2$NumWordsRemoved = rowSums(as.matrix(dtmRemoved))**bold**
What is the average number of words added?

```{r}
wikiTrain3 = subset(wikiWords2, spl==TRUE)
wikiTest3 = subset(wikiWords2, spl==FALSE)
wikiCART3 = rpart(Vandal~., data=wikiTrain3, method="class")

# Out-of-Sample Performnce of the Model (Make predictions on the wikiTest3 set)
testPredictCART3= predict(wikiCART3, newdata=wikiTest3,type="class") #if you add the argument type="class" when making predictions, the output of predict will automatically use a threshold of 0.5.

#We are interested in the accuracy of our model on the test set, i.e. out-of-sample. 
#First we compute the confusion matrix
table(wikiTest3$Vandal,testPredictCART3)
```
> (514+248)/(514+104+297+428) = 0.5689911

######Problem 3.1 - Using Non-Textual Data
```{r}
wikiWords3 = wikiWords2

wikiWords3$Minor = wiki$Minor
wikiWords3$Loggedin = wiki$Loggedin

wikiTrain4 = subset(wikiWords3, spl==TRUE)
wikiTest4 = subset(wikiWords3, spl==FALSE)
wikiCART4 = rpart(Vandal~., data=wikiTrain4, method="class") 

# Out-of-Sample Performnce of the Model (Make predictions on the test set wikiTest4)
testPredictCART4= predict(wikiCART4, newdata=wikiTest4,type="class") 

#We are interested in the accuracy of our model on the test set, i.e. out-of-sample. 
#First we compute the confusion matrix
table(wikiTest4$Vandal,testPredictCART4) 

```
>(595+241)/(595+23+304+241) = 0.7188306.

_______________________________________________________________________________________________________________

#Automating Reviews in Medicine

##### problem 1.1
```{r}
trials = read.csv("Downloads/clinical_trial.csv", stringsAsFactors=FALSE),

summary(nchar(trials$abstract))

```
> answer :3708

##### problem 1.2
```{r}
table(nchar(trials$abstract) == 0)
```
> answer : 112

##### problem 1.3

```{r}
trials$title[which.min(nchar(trials$title))]
```
> A decade of letrozole: FACE


##### problem 2.1
Because we have both title and abstract information for trials, we need to build two corpora instead of one. Name them corpusTitle and corpusAbstract.

Following the commands from lecture, perform the following tasks (you might need to load the "tm" package first if it isn't already loaded). Make sure to perform them in this order.

1) Convert the title variable to corpusTitle and the abstract variable to corpusAbstract.

2) Convert corpusTitle and corpusAbstract to lowercase.

3) Remove the punctuation in corpusTitle and corpusAbstract.

4) Remove the English language stop words from corpusTitle and corpusAbstract.

5) Stem the words in corpusTitle and corpusAbstract (each stemming might take a few minutes).

6) Build a document term matrix called dtmTitle from corpusTitle and dtmAbstract from corpusAbstract.

7) Limit dtmTitle and dtmAbstract to terms with sparseness of at most 95% (aka terms that appear in at least 5% of documents).

8) Convert dtmTitle and dtmAbstract to data frames (keep the names dtmTitle and dtmAbstract).

If the code length(stopwords("english")) does not return 174 for you, then please run the line of code in this file, which will store the standard stop words in a variable called sw. When removing stop words, use tm_map(corpusTitle, removeWords, sw) and tm_map(corpusAbstract, removeWords, sw) instead of tm_map(corpusTitle, removeWords, stopwords("english")) and tm_map(corpusAbstract, removeWords, stopwords("english")).

How many terms remain in dtmTitle after removing sparse terms (aka how many columns does it have)?
```{r}
corpusTitle= Corpus(VectorSource(trials$title))
corpusAbstract= Corpus(VectorSource(trials$abstract))

corpusTitle=tm_map(corpusTitle, content_transformer(tolower))
corpusAbstract=tm_map(corpusAbstract, content_transformer(tolower))

corpusTitle=tm_map(corpusTitle, removePunctuation)
corpusAbstract=tm_map(corpusAbstract, removePunctuation)

corpusTitle=tm_map(corpusTitle, removeWords, stopwords("english"))
corpusAbstract=tm_map(corpusAbstract, removeWords, stopwords("english"))

corpusTitle=tm_map(corpusTitle, stemDocument)
corpusAbstract=tm_map(corpusAbstract, stemDocument)

Corpus = tm_map(corpusTitle, PlainTextDocument)
Corpus = tm_map(corpusAbstract, PlainTextDocument)

dtmTitle= DocumentTermMatrix(corpusTitle)
dtmAbstract= DocumentTermMatrix(corpusAbstract)

dtmTitle= removeSparseTerms(dtmTitle, 0.95)
dtmAbstract= removeSparseTerms(dtmAbstract, 0.95)
dtmAbstract
dtmTitle
```

##### Problem 3.2 - Preparing the Corpus
What is the most frequent word stem across all the abstracts? Hint: you can use colSums() to compute the frequency of a word across all the abstracts.
```{r}
csAbstract = colSums(dtmAbstract)

which.max(csAbstract)
```
> 366

##### Problem 3.2 - Building a Model
Now that we have prepared our data frame, it's time to split it into a training and testing set and to build regression models. Set the random seed to 144 and use the sample.split function from the caTools package to split dtm into data frames named "train" and "test", putting 70% of the data in the training set.

What is the accuracy of the baseline model on the training set? (Remember that the baseline model predicts the most frequent outcome in the training set for all observations.)


```{r}
set.seed(144)

spl =  sample.split(dtm$trial, 0.7)
train = subset(dtm, spl == TRUE)
test = subset(dtm , spl == FALSE)

#accuracy baseline model
table(train$trial)
```
>730/(730+572)=0.5606759

#####PROBLEM 3.4 - BUILDING A MODEL

library(rpart)
library(rpart.plot)

#Let's  use CART to build a predictive model, using the rpart() function to predict responsive using all of the other variables as our independent variables and the data set train.
```{r}
trialCART= rpart(trial~., data=train, method="class") #the method="class" argument as this is a classification problem
```
#Plotting the CART model
```{r}
prp(trialCART)
```
>Tphase

#####Problem 3.5 - Building a Model
Obtain the training set predictions for the model (do not yet predict on the test set). Extract the predicted probability of a result being a trial (recall that this involves not setting a type argument, and keeping only the second column of the predict output). What is the maximum predicted probability for any result?
```{r}
predTrain= predict(trialCART)
max(predTrain[,2])
```