---
title: "R"
output: html_notebook
---
# Investment Management Under Taxation

Billions of shares of stock, or fractions of ownership in a business, are traded on the stock market every day. Over half of all adults in the United States own stocks and 1.2 billion people worldwide invest in the stock market. Many people invest in stocks to increase their wealth and to increase their earnings beyond their salary. If the business that you own stock in does well, then your stock value will increase and you will make money.

An individual who owns stock can sell their shares, or a fraction of their shares, to get cash that can be used for a down payment on a home, to buy a new car, or for any other purchase. However, when you sell stock, you have to pay both a transaction fee and  tax on the money you gain. If you own many different stocks, you have to decide what stocks and how much to sell to make sure you have enough cash for your purchase. In this problem, we'll use linear optimization to decide which shares of stock and how many you need to sell in order to have enough cash to make your purchase, and to maintain a strong portfolio of stocks. 

___________________________________________________________________________________________________________________________________________________
##### Problem 1.1 - Formulating the Problem
Suppose that, last year, you purchased 150 shares of eight different stocks (for a total of 1200 shares). The spreadsheet Investment.ods for LibreOffice or OpenOffice, and Investment.xlsx for Microsoft Excel, lists the stocks that you purchased, the price you purchased them for last year, the current price, and the price estimate for next year.

You would like to sell enough shares of stock today to generate $10,000 to use as part of a down payment on a new home. You need to decide how many shares of which stocks to sell in order to generate $10,000, after taxes and transaction costs, while maximizing the estimated value of your stock portfolio next year. Let's formulate this as a linear optimization problem.

**How many decision variables should your model have?**
```{r}
8 
```
> We need one decision variable for each stock, representing the number of shares to sell of that stock. Since we have 8 stocks, there are 8 decision variables.

##### Problem 1.2 - Formulating the Problem
*We'll assume for this problem that you can't sell more shares of stock than you own, and you can't buy additional shares. What is the maximum value your decision variables can be?*
```{r}
150
```
> You can't sell more shares that you own, and since you own 150 shares of each stock, the decision variables can't be any larger than 150.

*What is the minimum value your decision variables can be?*
```{r}
0
```
> Since you can't buy additional shares (giving the decision variables negative values) the minimum value the decision variables can be is 0.

##### Problem 1.3 - Formulating the Problem 
Your objective is to maximize the estimated value of your stock portfolio next year. To do this, you should sum the estimated value of each stock next year. Suppose you sell  ð‘¥  shares of your stock in Microsoft. What is the estimated value of your Microsoft stock next year?
```{r}
34.55 * (150 - x )
```

> Microsoft stock to be worth $34.55 next year. Since you will have (150 - x) shares next year (after selling x shares), the estimated value of your stock is $34.55(150 - x).

##### Problem 1.4 - Formulating the Problem
You need to make sure you get $10,000 in cash from selling your stocks, after taxes and transaction costs. How much would you get in cash, after taxes and transaction costs, if you sell 50 shares of your Intel stock?

```{r}
purchaseprice  = 22.54 * 50
currentprice =  23.67 * 50
gainsaftertax = (currentprice - purchaseprice) * 0.3
transactioncost = currentprice * 0.01
netcashflow = currentprice - gainsaftertax - transactioncost
```
> net cash flow = 1154.715

##### Problem 2.1 - Analyzing the Solution
In the optimal solution, which stocks do you sell at least one share of? Select all that apply.

> Intel, Yahoo, Microsoft, look into excel file

##### Problem 2.2 - Analyzing the Solution
What is the objective value of the optimal solution (the estimated value of your portfolio of stocks next year)?

> objective value: 26773.66271 

##### Problem 2.3 - Analyzing the Solution
How many shares of stock in total should you sell to make sure you have enough cash, according to the optimal solution? (Assume that you can sell fractional shares.)

> 367.723 = 67.723 Yahoo shares + 150 Microsoft shares + 150 Intel Shares

##### Problem 3.1 - Adjusting the Formulation
As an invester, you like having a portfolio of eight different stocks because it diversifies your investment. If one or two stocks do poorly this year, you won't worry as much because you have many other stocks. In the optimal solution for this problem, you sold all of your shares of some stocks, but you would like to keep at least half of the shares of each of your stocks.
Adjust the formulation so that you sell no more than 75 shares of each stock, and solve it again.
*In the optimal solution, you sell at least one share of which of your stocks? Select all that apply.*
> Yahoo!, GE, Microsoft, Cisco, Intel, Pfizer, change upperbound 150 > 75

##### Problem 3.2 - Adjusting the Formulation
> Objective value: 26468.54116


##### Problem 3.3 - Adjusting the Formulation
Even though your investment is worth a bit less next year by diversifying, you expect this diversification to help you long term.
However, you notice that you expect the Yahoo! stock to decrease in value next year. So, while you would like to sell no more than 75 shares of your other stocks, you would like to sell exactly 100 shares of your Yahoo! stock. Adjust your formulation in LibreOffice again, and re-solve to get the new optimal solution.
*You now sell at least one share of how many different stocks?*

> 5:Yahoo!, General Electric, Microsoft, Intel, and Pfizer; In excel add decision boundary , ahoo  =  sell 100 stocks*

__________________________________________________________________________________________________________________________________________________
# filatoi riuniti
The northern regions of Italy are the heartlands of the Italian textile industry, providing textile products for many great Italian (and non-Italian) fashion houses. Most of today's companies were founded at the turn of the century and have grown and thrived despite a series of twentieth century catastrophes, including the Great Depression and World War II. Sales in the entire Italian textile and apparel industry were about $80 billion in 2007. The Italian textile-fashion industry employs over 400,000 people. 

Filatoi Riuniti is a family-owned spinning mill located in northwest Italy. They produce cotton yarn, which is one of the raw materials that is used to produce the fabrics that are then cut and sewn by Italian stylists into the worldwide famous Italian clothes. 

Demand for Filatoi Riuniti's production is strong, but their spinning machine capacity is insufficient to meet its production orders. They decided to outsource part of the spinning production to six local family-owned spinning mills: Ambrosi, Bresciani, Castri, De Blasi, Estensi, and Giuliani. The local mills charge higher prices for spinning finer yarns, so Filatoi Riuniti has decided to spin as much as possible of the finer yarns entirely in-house and to outsource only the spinning of low-end (coarser) yarns. Last month, they faced a total demand of 104,500 kg of cotton and they outsourced 32,000 kg of the low-end sizes. 

Filatoi Riuniti hired a consulting firm to see if the outsourcing strategies to the six local mills could be improved. After analyzing the data, they immediately saw the potential for very large savings through more optimal outsourcing strategies. In this problem, we'll improve the outsourcing strategy of Filatoi Riuniti using linear optimization. (Note: This problem is based off of a real case, but the names have been changed for privacy reasons.)

##### Problem 1.1 - Formulating the Optimization Problem

Filatoi Riuniti produces four different sizes of yarn: extra fine, fine, medium, and coarse. Their current strategy is to produce all of the fine and extra fine yarn in-house, outsource some of the medium yarn to the lowest-price mill, and outsource all of the coarse yarn to the other mills. To try to improve this strategy, let's set up an optimization problem.

The decision variables are the amount of yarn of each size that each company should be assigned to produce. There are four sizes, and seven companies (including Filatoi Riuniti). These decision variables are set up in the spreadsheet. Note that the decision variables should be non-negative - none of the mills can produce a negative amount of yarn!

The objective is to minimize costs. The prices charged by the six local mills for production, as well as the production cost at Filatoi Riuniti, are provided in the spreadsheet. Additionally, the yarn that is spun by the six local mills needs to be transported. The transportation costs per kg of yarn are also provided in the spreadsheet.

Our model has two types of constraints: capacity constraints and demand constraints. The spreadsheet contains a table showing the production capacity and production rate per product for each of the mills (including Filatoi Riuniti). For example, at the Bresciani mill, it takes 0.70 hours to produce one kg of extra fine yarn, and there are a total of 3,000 machines hours per month available at this mill. There is also a table that estimates the demand for the four yarn sizes in the current month. We should produce at least the demand for each yarn type.

Formulate and solve this problem in LibreOffice (or in the spreadsheet software you are using). The objective should minimize the sum of the production and transportation costs. You should have the following constraints: non-negative decision variables, a capacity constraint for each mill, and a demand constraint for each type of yarn. Remember that Ambrosi and De Blasi can't produce extra fine yarn, so you should also account for this.
*What is the objective value of the solution?*

> The objective solution : 1382544.334

##### Problem 1.2 - Formulating the Optimization Problem
How many kg of medium yarn is outsourced?

> 9182.982595

##### Problem 1.3 - Formulating the Optimization Problem
How many kg of fine yarn is outsourced?

> 6250, We outsource 6250 kg of fine yarn to Ambrosi, and produce the rest in-house. So a total of 6250 kg of fine yarn is outsourced.

##### Problem 1.4 - Formulating the Optimization Problem
At which mills does Filatoi use the maximum capacity available? Select all that apply.

> Ambrosi, Bresciani, Castri, Esteni, Filatoi Riuniti, Giuliani

##### Problem 2.1 - Sensitivity Analysis

In Problems 2 and 3, we'll be checking how sensitive our model is to the data and constraints. If you change the constraints or data values to answer any of the questions, make sure to change them back to the original values before continuing on to the next problem.

Filatoi Riuniti should obviously consider increasing its spinning machine capacity. They could slightly expand the production capacity of the existing machines by renting an upgrade. This would increase their spinning production capacity by 600 hours/month. The monthly rental cost is $1,500/month. Would you recommend that they rent the upgrade?

> If Filattoi Riuniti increase spinning capacity by 600 hours/month change in new new objective value will be = (1381273,7-1382544.334) = -1270.5, an improvement in saving costs. upgrade cost $1500 month, so not worth it

##### Problem 2.2 - Sensitivity Analysis
Alternatively, Filatoi Riuniti could increase its spinning machine capacity by renting another spinning machine for the production of only medium size yarn, for a monthly rental cost of 3,000. The machine has a production capacity of 300 hours per month (the machine would run at the same rate of 0.425 hours/Kg). Suppose that the estimated production cost of running this machine is less than that for Filatoi Riuniti's existing machines and is estimated to be 5.70/Kg (as opposed to $11.40/Kg for their existing machines). Would you recommend that Filatoi Riuniti rent the machine?

>Yes they should rent the machine 

###### Problem 2.3 - Sensitivity Analysis
Suppose that the shadow price of the medium yarn demand constraint is 12.30, with an allowable increase of 5388. Recall that the shadow price is the amount that the objective increases per unit increase in the right hand side of the constraint. So a shadow price of 12.30 for the medium yarn demand constraint means that the cost increases by $12.30 for every unit increase in the medium yarn demand. The allowable increase of 5388 means that this shadow price hold up to an increase of 5388 in the right hand side of this constraint. If we increase the right hand side by more than 5388, the shadow price will change, and it is impossible to know how it changes without re-solving the model.

A new client is interested in purchasing up to 5,000 kg/month of medium size yarn. What is the minimum price per kg of yarn that Filatoi Riuniti should quote to this new client? (In answering this question, assume that Filatoi has not decided to increase its spinning machine capacity, and that Filatoi does not want to change the prices that they currently charge their existing clients.)

> (X) Filatoi should charge at least $12.30 per kg to the new client, since this is the amount it will cost per-unit to increase the medium yarn demand by 5,000.

###### Problem 2.4 - Sensitivity Analysis
Now suppose that the client wants to purchase 6,000 kg/month of medium size yarn. Now what is the minimum price per kg of yarn that Filatoi Riuniti should quote to this new client? (In answering this question, assume that Filatoi has not decided to increase its spinning machine capacity, and that Filatoi does not want to change the prices that they currently charge their existing clients.)

>12.45


#### Problem 3.1 - Data Estimates
In many situations, the problem data is estimated but is not known for sure. In this problem, the optimization model is based in part on the prices charged by the local mills to Filatoi Riuniti and on an estimate of Filatoi Riuniti's internal production costs. The plant manager, the accounting department, and you estimate that Filatoi Riuniti's internal production costs could vary within a 5% range of the numbers given.
If Filatoi Riuniti's production costs increase by 5%, by how much does the objective value of the solution change? You should adjust and re-optimize the model to answer this question.

> 48924.5

##### Problem 3.2 - Data Estimates
If Filatoi Riuniti's production costs decrease by 5%, by how much does the objective value of the solution change (in absolute difference)? You should adjust and re-optimize the model to answer this question.

> 48924.5

##### Problem 3.3 - Data Estimates
Now you think that the production capacity of one of your local mills, De Blasi, could vary within a 20% range of the estimated value. In the current solution, De Blasi produces 2040.125 kg of medium yarn, and none of the other types of yarn. If De Blasi's capacity is actually 20% higher or lower than the estimated value, will the solution change? HINT: Think about De Blasi's capacity constraint.

> No,De Blasi's capacity constraint is not binding in the current solution. While they have a capacity of 2600 hours, we only use 714 hours. So even if the capacity changes by 20%, the solution will not change.


__________________________________________________________________________________________________________________________________________________
# Gasoline Blending

One of the earliest and most successful applications of linear optimization was in the oil industry in the 1950s. According to Bill Drew, the former manager of research for Exxon Mobil, the company used linear optimization to "schedule our tanker fleets, design port facilities, blend gasoline, create financial models, you name it."  In this problem, we'll use optimization to blend gasoline. (While the application of this problem is real, the data that we will be using here has been created for this problem.) 

Gasoline blending occurs in oil refineries, where crude oil is processed and refined into more useful products, such as gasoline and diesel fuel. We will consider three products: super gasoline, regular gasoline, and diesel fuel. These can be made by mixing three different types of crude oil: crude 1, crude 2, and crude 3. Each product is distinguished by its octane rating, which measures the quality of the fuel, and its iron content, which is a contaminant in the gas. The crude oils each have an octane rating and iron content as well. The following table shows the required octane ratings and iron contents for each of the products, as well as the known octane ratings and iron contents of each of the crude oils:

Product or Oil	Octane Rating	Iron Content
Super Gasoline	at least 10	no more than 1
Regular Gasoline	at least 8	no more than 2
Diesel Fuel	at least 6	no more than 1
Crude 1	12	0.5
Crude 2	6	2.0
Crude 3	8	3.0
The gasoline produced must meet these standards for octane ratings and iron content. The octane rating and iron content of a product is the weighted average of the octane rating and iron content of the crude oils used to produce it. For example, if we produce regular gasoline using 20 barrels of Crude 1, 5 barrels of Crude 2, and 10 barrels of Crude 3, the Octane Rating of the regular gasoline would be:

(20*12 + 5*6 + 10*8)/35 = 10

The numerator is the number of barrels of Crude 1 used times the octane rating of Crude 1, plus the number of barrels of Crude 2 used times the octane rating of Crude 2, plus the number of barrels of Crude 3 used times the octane rating of Crude 3. The denominator is the total number of barrels used.

Similarly, the iron content of the regular gasoline would be:

**(20*0.5 + 5*2.0 + 10*3.0)/35 = 1.43**

The objective of the oil company is to maximize profit. The following table gives the sales price (revenue) for one barrel of each of the products: 

Product	Sales Price      
Super Gasoline  	$70
Regular Gasoline	$60
Diesel Fuel	      $50

Oil	Purchase Price
Crude 1	$45
Crude 2	$35
Crude 3	$25
We would like to maximize the amount made by selling the products, minus the amount it costs to buy the crude oils.

The company can only buy 5,000 barrels of each type of crude oil, and can process no more than 14,000 barrels total of crude oil. One barrel of crude oil makes one barrel of gasoline or fuel (nothing is lost in the conversion).

12ð‘¥+6ð‘¦+8ð‘§ð‘¥+ð‘¦+ð‘§â‰¥10 

It needs to be formulated as:

12ð‘¥+6ð‘¦+8ð‘§â‰¥10(ð‘¥+ð‘¦+ð‘§) 

##### Problem 1.1 - The Formulation
How many decision variables are there in this optimization model?

> 9, The are nine decision variables in this optimization model - one for each product and each type of crude oil. So we should have one decision variable for the amount of Crude 1 to use in Super Gasoline, one decision variable for the amount of Crude 2 to use in Super Gasoline, etc.

##### Problem 1.2 - The Formulation
How many constraints are in the model, not including the non-negativity constraints?

> 10, There are 10 constraints in this model. Three for the octane limits, three for the lead limits, three for each of the crude oil limits, and one for the total crude limit.

##### Problem 2.1 - The Solution
Solve your optimization model in LibreOffice (or in the spreadsheet software you are using). What is the objective value of the solution?

> Objective Value 375000, look into the excel file for the answer.

##### Problem 2.2 - The Solution
In the optimal solution, we produce the MOST barrels of which gasoline?

> Regular Gasoline, The total number of barrels of each type of gasoline produced can be computed by summing the total number of barrels of crude oil used to produce each type of gasoline. We produce far more barrels of regular gasoline than either of the other two types in the optimal solution.

##### Problem 2.3 - The Solution
In the optimal solution, how many barrels of diesel fuel are produced?

> 0, No barrels of crude oil are used to produce diesel fuel, so 0 barrels of diesel fuel are produced.

##### Problem 2.4 - The Solution
The company is unhappy with this solution, because they are exceeding the customer demand for regular gasoline. They estimate that the customer demand for super gasoline is 3,000 barrels, the customer demand for regular gasoline is 2,000 barrels, and the customer demand for diesel fuel is 1,000 barrels. They don't want to produce more than the customer demand, since they will lose revenue. Add constraints to your model to make sure that the solution produces no more than the customer demand for each of the products, and re-solve your model.
*What is the objective value now?*

> The new objective value : 150,000

##### Problem 2.5 - The Solution
Which of the crude oils are used in the solution? Select all that apply.

> Crude 1 and crude 3

##### Problem 3.1 - Sensitivity Analysis and Shadow Prices
The following are the shadow prices for each of the demand constraints:
Super gasoline demand shadow price = 29
Regular gasoline demand shadow price = 27
Diesel fuel demand shadow price = 9

The super gasoline shadow price holds to an allowable increase of 1250, the regular gasoline shadow price holds to an allowable increase of 2500, and the diesel fuel shadow price holds to an allowable increase of 1250. The "allowable increase" is the amount that you can increase the right-hand side and still use this shadow price. For a larger increase, the shadow price will change, and you have to resolve the model to get the new shadow price.
Please answer the following questions without re-solving the optimization model.
*What does a shadow price of 29 for the super gasoline demand constraint mean?*

> For one additional barrel of demand of super gasoline, the total profit will increase by 29.
 
##### Problem 3.1 - Sensitivity Analysis and Shadow Prices
According to the shadow prices, which type of gasoline should the company advertise to increase demand? Suppose that advertising costs $2 per unit increase in demand regardless of the type of gasoline being advertised. (Also suppose the company can only choose one type of gasoline to advertise.)

> For one additional barrel of demand of super gasoline, the total profit will increase by 29.

##### Problem 3.2 - Sensitivity Analysis and Shadow Prices
According to the shadow prices, which type of gasoline should the company advertise to increase demand? Suppose that advertising costs $2 per unit increase in demand regardless of the type of gasoline being advertised. (Also suppose the company can only choose one type of gasoline to advertise.)

> Super Gasoline, The gasoline with the largest shadow price is the most profitable if demand is increased, since an extra unit of demand increases profits the most.

##### Problem 3.3 - Sensitivity Analysis and Shadow Prices
How much additional profit would you gain if the super gasoline demand increased by 500?
(You can assume that the demand increased on its own and you didn't have to pay anything in advertising for this increase).

> Shadow price = 29, Demand = 500  = 29*500 =14.500. Since the shadow price of super gasoline demand is 29, and this holds up to an allowable increase of 1250

##### Problem 3.4 - Sensitivity Analysis and Shadow Prices
How much additional profit would you gain if you increased the super gasoline demand by 1500?
(You can assume that the demand increased on its own and you didn't have to pay anything in advertising for this increase).

>Can not be determined from the given information without re-solving the optimization model, since the shadow price only holds up to an allowable increase of 1250, we can't answer this question given the information provided.






