---
title: "unit 6"
output: html_notebook
---
# Document clustering 

Document clustering, or text clustering, is a very popular application of clustering algorithms. A web search engine, like Google, often returns thousands of results for a simple query. For example, if you type the search term "jaguar" into Google, around 200 million results are returned. This makes it very difficult to browse or find relevant information, especially if the search term has multiple meanings. If we search for "jaguar", we might be looking for information about the animal, the car, or the Jacksonville Jaguars football team. 

Clustering methods can be used to automatically group search results into categories, making it easier to find relevant results. This method is used in the search engines PolyMeta and Helioid, as well as on FirstGov.gov, the official Web portal for the U.S. government. The two most common algorithms used for document clustering are Hierarchical and k-means. 

In this problem, we'll be clustering articles published on Daily Kos, an American political blog that publishes news and opinion articles written from a progressive point of view. Daily Kos was founded by Markos Moulitsas in 2002, and as of September 2014, the site had an average weekday traffic of hundreds of thousands of visits. 

The file dailykos.csv contains data on 3,430 news articles or blogs that have been posted on Daily Kos. These articles were posted in 2004, leading up to the United States Presidential Election. The leading candidates were incumbent President George W. Bush (republican) and John Kerry (democratic). Foreign policy was a dominant topic of the election, specifically, the 2003 invasion of Iraq. 

Each of the variables in the dataset is a word that has appeared in at least 50 different articles (1,545 words in total). The set of  words has been trimmed according to some of the techniques covered in the previous week on text analytics (punctuation has been removed, and stop words have been removed). For each document, the variable values are the number of times that word appeared in the document.

______________________________________________________________________________________________________________________

###### Problem 1.1 - Hierarchical Clustering
Let's start by building a hierarchical clustering model. First, read the data set into R. Then, compute the distances (using method="euclidean"), and use hclust to build the model (using method="ward.D"). You should cluster on all of the variables.
Running the dist function will probably take you a while. Why? Select all that apply.
```{r}
dailykos = read.csv("Documents/Moocs/MIT AE/dailykos.csv")
str(dailykos)
dim(dailykos)
#compute the distance
distance = dist(dailykos, method = "euclidean")
#hierarchical clustering of dailykos
DKHC=hclust(distance, method="ward.D")
plot(DKHC)
```
>
1. We have a lot of observations, so it takes a long time to compute the distance between each pair of observations.
2. We have a lot of variables, so the distance computation is long.

###### Problem 1.2 - Hierarchical Clustering

Plot the dendrogram of your hierarchical clustering model. Just looking at the dendrogram, which of the following seem like good choices for the number of clusters? Select all that apply.

```{r}
rect.hclust(DKHC, k = 3, border = "red")
Clusters = cutree(DKHC, k = 3)
Clusters
```
>The choices 2 and 3 are good cluster choices according to the dendrogram, because there is a lot of space between the horizontal lines in the dendrogram in those cut off spots (draw a horizontal line across the dendrogram where it crosses 2 or 3 vertical lines). The choices of 5 and 6 do not seem good according to the dendrogram because there is very little space

###### Problem 1.3 - Hierarchical Clustering

In this problem, we are trying to cluster news articles or blog posts into groups. This can be used to show readers categories to choose from when trying to decide what to read. Just thinking about this application, what are good choices for the number of clusters? Select all that apply.
> Thinking about the application, it is probably better to show the reader more categories than 2 or 3. These categories would probably be too broad to be useful. Seven or eight categories seems more reasonable.

###### Problem 1.4 - Hierarchical Clustering
Let's pick 7 clusters. This number is reasonable according to the dendrogram, and also seems reasonable for the application. Use the cutree function to split your data into 7 clusters.
Now, we don't really want to run tapply on every single variable when we have over 1,000 different variables. Let's instead use the subset function to subset our data by cluster. Create 7 new datasets, each containing the observations from one of the clusters.
1. How many observations are in cluster 3?
2. Which cluster has the most observations?
```{r}
DKclusters = cutree(DKHC, k = 7)
 C1 = subset(dailykos, DKclusters == 1)
 C2 = subset(dailykos, DKclusters == 2)
 C3 = subset(dailykos, DKclusters == 3)
 C4 = subset(dailykos, DKclusters == 4)
 C5 = subset(dailykos, DKclusters == 5)
 C6 = subset(dailykos, DKclusters == 6)
 C7 = subset(dailykos, DKclusters == 7)
table(DKclusters)
```
###### Problem 1.5 - Hierarchical Clustering
Instead of looking at the average value in each variable individually, we'll just look at the top 6 words in each cluster. To do this for cluster 1, type the following in your R console (where "HierCluster1" should be replaced with the name of your first cluster subset):
**tail(sort(colMeans(HierCluster1)))**
This computes the mean frequency values of each of the words in cluster 1, and then outputs the 6 words that occur the most frequently. The colMeans function computes the column (word) means, the sort function orders the words in increasing order of the mean values, and the tail function outputs the last 6 words listed, which are the ones with the largest column means.
What is the most frequent word in this cluster, in terms of average value? Enter the word exactly how you see it in the output:
```{r}
mostfreqC1 = tail(sort(colMeans(C1)))
mostfreqC1
```
> bush 1.7053712

###### Problem 1.6 - Hierarchical Clustering
Now repeat the command given in the previous problem for each of the other clusters, and answer the following questions.
```{r}
mostfreqC2 = tail(sort(colMeans(C2)))
mostfreqC3 = tail(sort(colMeans(C3)))
mostfreqC4 = tail(sort(colMeans(C4)))
mostfreqC5 = tail(sort(colMeans(C5)))
mostfreqC6 = tail(sort(colMeans(C6)))
mostfreqC7 = tail(sort(colMeans(C7)))

#split()  function. Given a vector assigning groups like hierGroups, you could split dailykos
HierCluster = split(dailykos, DKclusters)
#using lapply() rather than doing clusterwise
topwords = lapply(HierCluster, function(c) tail(sort(colMeans(c))))
topwords
```
>
**Which words best describe cluster 2?** Answer: november, poll, vote, challenge
**Which cluster could best be described as the cluster related to the Iraq war?** Answer: Cluster 5
**John Edwards with the running mate of John Kerry. Given this information, which cluster best corresponds to the democratic party?** Answer: Cluster 7

###### Problem 2.1 - K-Means Clustering
Run k-means clustering, setting the seed to 1000 right before you run the kmeans function. Again, pick the number of clusters equal to 7. You don't need to add the iters.max argument.
Subset your data into the 7 clusters (7 new datasets) by using the "cluster" variable of your kmeans output.
```{r}
set.seed(1000)
k=7
KMC = kmeans(dailykos, centers = k)
str(KMC)
table(KMC$cluster)

kmc1 = subset(dailykos, KMC$cluster == 1)
kmc2 = subset(dailykos, KMC$cluster == 2)
kmc3 = subset(dailykos, KMC$cluster == 3)
kmc4 = subset(dailykos, KMC$cluster == 4)
kmc5 = subset(dailykos, KMC$cluster == 5)
kmc6 = subset(dailykos, KMC$cluster == 6)
kmc7 = subset(dailykos, KMC$cluster == 7)


```
>
**How many observations are in Cluster 3?** Answer : 300
**Which cluster has the most observations?** Answer: 4
**Which cluster has the fewest number of observations?** Answer: 5

###### Problem 2.2 - K-Means Clustering
Now, output the six most frequent words in each cluster, like we did in the previous problem, for each of the k-means clusters.
```{r}
tail(sort(colMeans(kmc1)))
tail(sort(colMeans(kmc2)))
tail(sort(colMeans(kmc3)))
tail(sort(colMeans(kmc4)))
tail(sort(colMeans(kmc5)))
tail(sort(colMeans(kmc6)))
tail(sort(colMeans(kmc7)))


KmeansClusterSplit = split(dailykos, KMC)
topwords = lapply(KmeansClusterSplit, function(c) tail(sort(colMeans(c))))
topwords
```
>
**Which k-means cluster best corresponds to the Iraq War?** Answer:3
**Which k-means cluster best corresponds to the democratic party? (Remember that we are looking for the names of the key democratic party leaders.)** Answer: 2

###### Problem 2.3 - K-Means Clustering
For the rest of this problem, we'll ask you to compare how observations were assigned to clusters in the two different methods. Use the table function to compare the cluster assignment of hierarchical clustering to the cluster assignment of k-means clustering.

Which Hierarchical Cluster best corresponds to K-Means Cluster 2?
```{r}
table(DKclusters, KMC$cluster)
```
> Hierarchical cluster 7 

###### - Problem 2.4 - K-Means Clustering
Which Hierarchical Cluster best corresponds to K-Means Cluster 3?
> cluster 5 

###### Problem 2.5 - K-Means Clustering
Which Hierarchical Cluster best corresponds to K-Means Cluster 7?
> No Hierarchical Cluster contains at least half of the points in K-Means Cluster 7.

###### Problem 2.6 - K-Means Clustering
Which Hierarchical Cluster best corresponds to K-Means Cluster 6?
> 


_____________________________________________________________________________________________________________________
# market Segmentation For Airlines

Market segmentation is a strategy that divides a broad target market of customers into smaller, more similar groups, and then designs a marketing strategy specifically for each group. Clustering is a common technique for market segmentation since it automatically finds similar groups given a data set. 
In this problem, we'll see how clustering can be used to find similar groups of customers who belong to an airline's frequent flyer program. The airline is trying to learn more about its customers so that it can target different customer segments with different types of mileage offers. 

*The file AirlinesCluster.csv contains information on 3,999 members of the frequent flyer program. This data comes from the textbook "Data Mining for Business Intelligence," by Galit Shmueli, Nitin R. Patel, and Peter C. Bruce.*

There are seven different variables in the dataset, described below:

**Balance** = number of miles eligible for award travel
**QualMiles** = number of miles qualifying for TopFlight status
**BonusMiles** = number of miles earned from non-flight bonus transactions in the past 12 months
**BonusTrans** = number of non-flight bonus transactions in the past 12 months
**FlightMiles** = number of flight miles in the past 12 months
**FlightTrans** = number of flight transactions in the past 12 months
**DaysSinceEnroll** = number of days since enrolled in the frequent flyer program

###### Problem 1.1 - Normalizing the Data
Looking at the summary of airlines, which TWO variables have (on average) the smallest values?
```{r}
airlines = read.csv("Documents/Moocs/MIT AE/airlinesCluster.csv")
summary(airlines)
colmeans(airlines)
``` 
>
**Looking at the summary of airlines, which TWO variables have (on average) the smallest values?** 
BonusMiles & FlighTrans
**Which TWO variables have (on average) the largest values?** 
Balance & BonusMiles

###### Problem 1.2 - Normalizing the Data
In this problem, we will normalize our data before we run the clustering algorithms. Why is it important to normalize the data before clustering?
> If we don't normalize the data, the clustering will be dominated by the variables that are on a larger scale.

###### Problem 1.3 - Normalizing the Data
In the normalized data, which variable has the largest maximum value?
```{r}
preproc = preProcess(airlines)
#normalize variables
airlinesNorm = predict(preproc, airlines)
round(colMeans(airlinesNorm),6)
sapply(airlinesNorm,sd)
apply(airlinesNorm, 2, max)
apply(airlinesNorm, 2, min)
```
**In the normalized data, which variable has the largest maximum value?** ANS: FlightMiles
**In the normalized data, which variable has the smallest minimum value?** ANS: DaysSinceEnroll
###### Problem 2.1 - Hierarchical Clustering
#Compute the distances between data points (using euclidean distance) and then run the Hierarchical clustering algorithm (using method="ward.D") on the normalized data. It may take a few minutes for the commands to finish since the dataset has a large number of observations for hierarchical clustering.
```{r}
distances = dist(airlinesNorm,method="euclidean")
airlineClust = hclust(distances,method="ward.D")
#plot the dendrogram of the hierarchical clustering process
plot(airlineClust)
```
> ANS : 6

###### Problem 2.2 - Hierarchical Clustering
Suppose that after looking at the dendrogram and discussing with the marketing department, the airline decides to proceed with 5 clusters. Divide the data points into 5 clusters by using the cutree function. How many data points are in Cluster 1?
```{r}
#Divide into 5 clusters
clusterGroups=cutree(airlineClust, k=5)
airlineClusters=split(airlinesNorm, clusterGroups)
table(clusterGroups)
```
>ANS : 776

###### Problem 2.3 - Hierarchical Clustering
Compared to the other clusters, Cluster 1 has the largest average values in which variables (if any)? Select all that apply.
```{r}
tapply(airlines$Balance, clusterGroups, mean)
tapply(airlines$QualMiles, clusterGroups, mean)
tapply(airlines$BonusMiles, clusterGroups, mean)
tapply(airlines$BonusTrans, clusterGroups, mean)
tapply(airlines$FlightMiles, clusterGroups, mean)
tapply(airlines$FlightTrans, clusterGroups, mean)
tapply(airlines$DaysSinceEnroll, clusterGroups, mean)

airlinesUnnormClusters<-split(airlines,clusterGroups)
round(sapply(airlinesUnnormClusters,colMeans),4)
```
> ANS : DaysSinceEnroll

###### Problem 2.4 - Hierarchical Clustering
Compared to the other clusters, Cluster 2 has the largest average values in which variables (if any)? Select all that apply.

> 
**Compared to the other clusters, Cluster 2 has the largest average values in which variables (if any)? Select all that apply.**Ans:QualMiles,FlightMiles,FlightTrans
**How would you describe the customers in Cluster 2?**
Ans: Customers who have accumulated a large amount of miles, and the ones with the largest number of flight transactions. 

###### Problem 2.5 - Hierarchical Clustering
>
**Compared to the other clusters, Cluster 3 has the largest average values in which variables (if any)? Select all that apply.** Ans:Balance,BonusMiles,BonusTrans
**How would you describe the customers in Cluster 3?**
Ans:Customers who have accumulated a large amount of miles, mostly through non-flight transactions.

###### Problem 2.6 - Hierarchical Clustering
>
**Compared to the other clusters, Cluster 4 has the largest average values in which variables (if any)? Select all that apply.** ANS: NONE
**How would you describe the customers in Cluster 4?**

###### Problem 2.7 - Hierarchical Clustering
>
**Compared to the other clusters, Cluster 5 has the largest average values in which variables (if any)? Select all that apply** ans: none
**How would you describe the customers in Cluster 5?** 
ans: Relatively new customers who don't use the airline very often.

###### Problem 3.1 - K-Means Clustering
Run the k-means clustering algorithm on the normalized data, again creating 5 clusters. Set the seed to 88 right before running the clustering algorithm, and set the argument iter.max to 1000.
```{r}
set.seed(88)
kmeansClust<- kmeans(airlinesNorm, centers=5, iter.max=1000)
table(kmeansClust$cluster)
```
> ans: 2 

###### Problem 3.2 - K-Means Clustering
Now, compare the cluster centroids to each other either by dividing the data points into groups and then using tapply, or by looking at the output of **kmeansClust$centers, where "kmeansClust" is the name of the output of the kmeans function. (Note that the output of kmeansClust$centers** will be for the normalized data. If you want to look at the average values for the unnormalized data, you need to use tapply like we did for hierarchical clustering.)

Do you expect Cluster 1 of the K-Means clustering output to necessarily be similar to Cluster 1 of the Hierarchical clustering output?
```{r}
airlinesUnnormClusters<-split(airlinesNorm,clusterGroups)
round(sapply(airlinesUnnormClusters,colMeans),4)
kmeansClust$centers
```
>No, because cluster ordering is not meaningful in either k-means clustering or hierarchical clustering. 

_____________________________________________________________________________________________________________________
# Predicting stock returns with cluster-then-predict

In the second lecture sequence this week, we heard about cluster-then-predict, a methodology in which you first cluster observations and then build cluster-specific prediction models. In the lecture sequence, we saw how this methodology helped improve the prediction of heart attack risk. In this assignment, we'll use cluster-then-predict to predict future stock prices using historical stock data.

When selecting which stocks to invest in, investors seek to obtain good future returns. In this problem, we will first use clustering to identify clusters of stocks that have similar returns over time. Then, we'll use logistic regression to predict whether or not the stocks will have positive future returns.

Each observation in the dataset is the monthly returns of a particular company in a particular year. The years included are 2000-2009. The companies are limited to tickers that were listed on the exchange for the entire period 2000-2009, and whose stock price never fell below $1. So, for example, one observation is for Yahoo in 2000, and another observation is for Yahoo in 2001. Our goal will be to predict whether or not the stock return in December will be positive, using the stock returns for the first 11 months of the year.

*For this problem, we'll use StocksCluster.csv, which contains monthly stock returns from the NASDAQ stock exchange. The NASDAQ is the second-largest stock exchange in the world, and it lists many technology companies. The stock price data used in this problem was obtained from infochimps, a website providing access to many datasets.*

This dataset contains the following variables:

* ReturnJan = the return for the company's stock during January (in the year of the observation). 
* ReturnFeb = the return for the company's stock during February (in the year of the observation). 
* ReturnMar = the return for the company's stock during March (in the year of the observation). 
* ReturnApr = the return for the company's stock during April (in the year of the observation). 
* ReturnMay = the return for the company's stock during May (in the year of the observation). 
* ReturnJune = the return for the company's stock during June (in the year of the observation). 
* ReturnJuly = the return for the company's stock during July (in the year of the observation). 
* ReturnAug = the return for the company's stock during August (in the year of the observation). 
* ReturnSep = the return for the company's stock during September (in the year of the observation). 
* ReturnOct = the return for the company's stock during October (in the year of the observation). 
* ReturnNov = the return for the company's stock during November (in the year of the observation). 
* PositiveDec = whether or not the company's stock had a positive return in December (in the year of the observation). 

This variable takes value 1 if the return was positive, and value 0 if the return was not positive.
For the first 11 variables, the value stored is a proportional change in stock value during that month. For instance, a value of 0.05 means the stock increased in value 5% during the month, while a value of -0.02 means the stock decreased in value 2% during the month.

###### Problem 1.1 - Exploring the Dataset
Load StocksCluster.csv into a data frame called "stocks". How many observations are in the dataset?
```{r}
stocks = read.csv("Documents/Moocs/MIT AE/StocksCluster.csv")
str(stocks)
```
>11580

###### Problem 1.2 - Exploring the Dataset
What proportion of the observations have positive returns in December?
```{r}
table(stocks$PositiveDec)
```
>6324

###### Problem 1.3 - Exploring the Dataset
What is the maximum correlation between any two return variables in the dataset? You should look at the pairwise correlations between ReturnJan, ReturnFeb, ReturnMar, ReturnApr, ReturnMay, ReturnJune, ReturnJuly, ReturnAug, ReturnSep, ReturnOct, and ReturnNov.
```{r}
correlation = cor(stocks)
#set >= to filter if TRUE or FALSE pairwaise correlations
kable(correlation >=0.15)
#look again to ReturnOct, ReturnNOV
```
>0.1916728

###### Problem 1.4 - Exploring the Dataset
```{r}
summary(stocks)
mean = colMeans(stocks[1:11])
which.max(mean)
which.min(mean)

```
>
**Which month (from January through November) has the largest mean return across all observations in the dataset?** ans: ReturnApr 4
**Which month (from January through November) has the smallest mean return across all observations in the dataset?** ans: ReturnSep 9

###### Problem 2.1 - Initial Logistic Regression Model
Run the following commands to split the data into a training set and testing set, putting 70% of the data in the training set and 30% of the data in the testing set.
use the stocksTrain data frame to train a logistic regression model (name it StocksModel) to predict PositiveDec using all the other variables as independent variables. Don't forget to add the argument family=binomial to your glm command.
What is the overall accuracy on the training set, using a threshold of 0.5?
```{r}
library(caTools)
set.seed(144)
spl = sample.split(stocks$PositiveDec, SplitRatio = 0.7)
stocksTrain = subset(stocks, spl == TRUE)
stocksTest = subset(stocks, spl == FALSE)
#train model
logmodel = glm(PositiveDec ~., data = stocksTrain, family = binomial)
#compute predictions on the training set
PredTrain = predict(logmodel, type="response")
#confusion matrix treshold > 0.5
table(stocksTrain$PositiveDec, PredTrain > 0.5)
#calculate accuracy by summing true positives and false negatives dividing by the total
```
>(990+3640)/(990+2689+787+3640) = 0.5711818

###### Problem 2.2 - Initial Logistic Regression Model
Now obtain test set predictions from StocksModel. What is the overall accuracy of the model on the test, again using a threshold of 0.5?
```{r}
# predictions of the Logistic Regression model, test set
PredTest = predict(logmodel, newdata = stocksTest, type = "response")
table(stocksTest$PositiveDec, PredTest > 0.5)
#calculate accuracy by summing true positives and false negatives.
```
>(417+1553)/(417+1160+344+1553) = 0.5670697

###### Problem 2.3 - Initial Logistic Regression Model
What is the accuracy on the test set of a baseline model that always predicts the most common outcome (PositiveDec = 1)?
```{r}
table(stocksTest$PositiveDec)
```
>1897/(1577+1897) = 0.5460564

###### Problem 3.1 - Clustering Stocks
Now, let's cluster the stocks. The first step in this process is to remove the dependent variable using the following commands:
```{r}
limitedTrain = stocksTrain
limitedTrain$PositiveDec = NULL
limitedTest = stocksTest
limitedTest$PositiveDec = NULL
```
>
**Why do we need to remove the dependent variable in the clustering phase of the cluster-then-predict methodology?**
Needing to know the dependent variable value to assign an observation to a cluster defeats the purpose of the methodology

###### Problem 3.2 - Clustering Stocks
In the market segmentation assignment in this week's homework, you were introduced to the preProcess command from the caret package, which normalizes variables by subtracting by the mean and dividing by the standard deviation.

In cases where we have a training and testing set, we'll want to normalize by the mean and standard deviation of the variables in the training set. 
* What is the mean of the ReturnJan variable in normTrain?
* What is the mean of the ReturnJan variable in normTest?
```{r}
library(caret)
preproc = preProcess(limitedTrain)
normTrain = predict(preproc, limitedTrain)
normTest = predict(preproc, limitedTest)
#mean Returns
mean(normTrain$ReturnJan)
mean(normTest$ReturnJan)
```
>
**What is the mean of the ReturnJan variable in normTrain?** ans: 2.100586e-17
**What is the mean of the ReturnJan variable in normTest?** ans: -0.0004185886

###### Problem 3.3 - Clustering Stocks
Why is the mean ReturnJan variable much closer to 0 in normTrain than in normTest?
>ans: The distribution of the ReturnJan variable is different in the training and testing set

###### Problem 3.4 - Clustering Stocks
Set the random seed to 144 (it is important to do this again, even though we did it earlier). Run k-means clustering with 3 clusters on normTrain, storing the result in an object called km.
Which cluster has the largest number of observations?
```{r}
set.seed(144)
km = kmeans(normTrain, centers = 3, iter.max = 1000)
table(km$cluster)
```
>ans: cluster 2

###### Problem 3.5 - Clustering Stocks
Recall from the recitation that we can use the flexclust package to obtain training set and testing set cluster assignments for our observations (note that the call to as.kcca may take a while to complete):
How many test-set observations were assigned to Cluster 2?
```{r}
library(flexclust)
km.kcca = as.kcca(km, normTrain)
clusterTrain = predict(km.kcca)
clusterTest = predict(km.kcca, newdata=normTest)
table(clusterTest)
```
> ans: 2080

###### Problem 4.1 - Cluster-Specific Predictions
Using the subset function, build data frames stocksTrain1, stocksTrain2, and stocksTrain3, containing the elements in the stocksTrain data frame assigned to clusters 1, 2, and 3, respectively (be careful to take subsets of stocksTrain, not of normTrain). Similarly build stocksTest1, stocksTest2, and stocksTest3 from the stocksTest data frame.
Which training set data frame has the highest average value of the dependent variable?
```{r}
stocksTrain1 = subset(stocksTrain, clusterTrain == 1)
stocksTrain2 = subset(stocksTrain, clusterTrain == 2)
stocksTrain3 = subset(stocksTrain, clusterTrain == 3)
stocksTest1 = subset(stocksTest, clusterTest == 1)
stocksTest2 = subset(stocksTest, clusterTest == 2)
stocksTest3 = subset(stocksTest, clusterTest == 3)

mean(stocksTrain1$PositiveDec)
mean(stocksTrain2$PositiveDec)
mean(stocksTrain3$PositiveDec)

#or
#stocksTrain11 <- split(stocksTrain, clusterTrain)
#stocksTest11 <- split(stocksTest, clusterTest)
#sapply(stocksTrain11, function(s){ mean(s$PositiveDec) })
```
>

###### Problem 4.2 - Cluster-Specific Predictions
Build logistic regression models StocksModel1, StocksModel2, and StocksModel3, which predict PositiveDec using all the other variables as independent variables. StocksModel1 should be trained on stocksTrain1, StocksModel2 should be trained on stocksTrain2, and StocksModel3 should be trained on stocksTrain3.
Which variables have a positive sign for the coefficient in at least one of StocksModel1, StocksModel2, and StocksModel3 and a negative sign for the coefficient in at least one of StocksModel1, StocksModel2, and StocksModel3? Select all that apply.
```{r}
StocksModel1 = glm(PositiveDec~., data=stocksTrain1, family=binomial)
StocksModel2 = glm(PositiveDec~., data=stocksTrain2, family=binomial)
Stocksmodel3 = glm(PositiveDec~., data=stocksTrain3, family=binomial)
summary(StocksModel1)
summary(StocksModel2)
summary(Stocksmodel3)

#stocksModels <- lapply(stocksTrain11, function(s){glm(s$PositiveDec ~ ., family=binomial, data=s)})
#sapply(stocksModels, function(m){ m$coefficients })

```
>

###### Problem 4.3 - Cluster-Specific Predictions
```{r}
PredictTest1 = predict(StocksModel1, newdata = stocksTest1, type = "response")
PredictTest2 = predict(StocksModel2, newdata = stocksTest2, type = "response")
PredictTest3 = predict(Stocksmodel3, newdata = stocksTest3, type = "response")

table(stocksTest1$PositiveDec, PredictTest1 > 0.5)
table(stocksTest2$PositiveDec, PredictTest2 > 0.5)
table(stocksTest3$PositiveDec, PredictTest3 > 0.5)
```
>
**What is the overall accuracy of StocksModel1 on the test set stocksTest1, using a threshold of 0.5?**
(30+774)/(30+471+23+774)=0.6194145
**What is the overall accuracy of StocksModel2 on the test set stocksTest2, using a threshold of 0.5?**
(388+757)/(388+626+309+757)=0.5504808
**What is the overall accuracy of StocksModel3 on the test set stocksTest3, using a threshold of 0.5?**
(49+13)/(49+13+21+13)=0.6458333

###### Problem 4.4 - Cluster-Specific Predictions
Using StocksModel1, make test-set predictions called PredictTest1 on the data frame stocksTest1. Using StocksModel2, make test-set predictions called PredictTest2 on the data frame stocksTest2. Using StocksModel3, make test-set predictions called PredictTest3 on the data frame stocksTest3.
```{r}
AllPredictions = c(PredictTest1, PredictTest2, PredictTest3)
AllOutcomes = c(stocksTest1$PositiveDec, stocksTest2$PositiveDec, stocksTest3$PositiveDec)
cmatoverall<-table(AllOutcomes, AllPredictions > 0.5)
cmatoverall
```
> 
**What is the overall test-set accuracy of the cluster-then-predict approach, again using a threshold of 0.5?**
(467 + 1544)/(467 + 1110 + 353 + 1544) = 0.5788716

We see a modest improvement over the original logistic regression model. Since predicting stock returns is a notoriously hard problem, this is a good increase in accuracy. By investing in stocks for which we are more confident that they will have positive returns (by selecting the ones with higher predicted probabilities), this cluster-then-predict model can give us an edge over the original logistic regression model.

